%!TeX root=../../main.tex
\chapter{Preface}                                 \label{ch:preface}
This thesis marks the end of a long academic journey that would not have been possible without the continuous support of all the people around me to whom I would like to dedicate this section.
\\[10pt]

First of all, I would like to thank my promotors and mentor for their support and guidance throughout this journey. Specifically, I would like to thank Professor E. Truyen for his understanding, patience and continuous aid in applying his expertise every step of the way. Naturally, I want to extend my gratitude to my mentor, G. Budigiri, as well who was an immense help in finding my way around the many pitfalls of this thesis and gladly answered all my questions no matter how big or small.
\\[10pt]

 I am forever grateful to my parents, whose love and support created the foundations of my academic pursuits. They gave me the opportunity to express myself and pursue this academic career, and I can say with certainty that this thesis would not have come to fruition without them. 
I also owe a debt of gratitude towards all my friends and family, whose words of encouragement in both good and challenging times never failed to motivate me.
Finally, I would like to thank my girlfriend for being my source of strength. In the moments when self-doubt crept in, she never ceased to stand by me with an unwavering belief in my capabilities, making her my guiding light through the rough patches of this journey.
\\[10pt]

Finally, I would like to express my heartfelt appreciation to each and every person who has contributed, in their own unique way, to the completion of this thesis. Their impact, however small, has left an unforgettable mark on this work.
\\[10pt]

Jasper Goris
\cleardoublepage
